---
title: "nlmixr Comparison"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{nlmixr comparison}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
## Examination of nlmixr estimation algorithm properties for richly sampled data and multiple models

The nlmixr/SAEM, and nlmixr/FOCEI parameter estimation algorithms were
compared with Monolix/SAEM and NONMEM/FOCEI in a richly-sampled data
setting with multiple models.

Richly sampled profiles were simulated for 4 different dose levels of
30 subjects each, for a range of test models with:

-	one- or two-compartmental disposition
-	oral (first-order absorption), intravenous (IV) bolus, or IV
     infusion administration
-	linear or Michaelis-Menten (MM) clearance.

In addition, three dosing and sampling scenarios were investigated: 
-	a single administration with 19 samples over 72 hours
-	seven repeated daily administrations, with 15 samples over 24
     hours after the 4th dose, 19 samples over 72 hours after the 7th
     dose, and 5 trough samples
-	the single administration profile followed by the repeated
     administrations profile, with a total of 58 samples over 12 days

These combinations provided a total of 36 test cases. Inter-individual
variability was applied to all pharmacokinetic parameters, and all
IIVs were set to 30% (implemented as a diagonal matrix with no
covariances). Proportional residual variability was set to 20%. All
one-compartment models had a population Vc of 70 L, and all
two-compartment models had an additional peripheral volume (Vp) of 40
L. For all oral absorption models, ka was set to 1.0 h-1. All models
with linear elimination had a CL of 4.0 L/h, and for all models with
non-linear MM elimination, CL was replaced with a Km of 250 mg/L and a
Vmax of 1000 mg/h. All two-compartment models had inter-compartmental
clearance (Q) set to 4.0 L/h.

```{r setup}

library(nlmixr)
library(dplyr)
library(xgxr)

## Function converting arguments to characters
Cs <- function(...){
    sapply(as.list(match.call())[-1], as.character)
}
expected_values <- list();
runno <- "run"

readData <- function(nlmixrVersion=c("1.1.1.3"), est=c("saem", "focei", "nlme"),
                     platform="unix",
                     runs=sort(Cs(U001,U002,U003,U004,U009,U010,U011,U012,U013,U014,
                                  U015,U020,U021,U022,U023,U024,U026,U025, U029, U030,U031,
                                  U032,U033,U035,U034,U040,U041,U042,U046,U047,U049,U048, 
                                  U054,U055,U056,U060,U061,U062,U063,U068, U069,U070))){
    .ret <- expand.grid(ver=nlmixrVersion, est=est, os=platform, run=runs, stringsAsFactors=FALSE) %>%
        mutate(src=paste0("values-", ver, "-", run,"_", est, "-", os, ".R"))
    .ret <- lapply(seq_along(.ret$ver), function(x){
        .c <- setNames(as.character(.ret[x, ]), names(.ret))
        .f <- file.path(system.file(package="nlmixr.examples"), "models", .c["src"]);
        if (file.exists(.f)){
            source(.f)
            .ret <- expected_values[["run"]]
            .ret1 <- setNames(.ret$parFixedDf$`Back-transformed`,
                       sapply(rownames(.ret$parFixedDf),
                              function(x){
                           if (x == "prop.err"){
                               x
                           } else {
                               substr(x, 2, nchar(x))
                           }
                       }))
            .ret2 <- setNames(.ret$parFixedDf$`SE`,
                       sapply(rownames(.ret$parFixedDf),
                              function(x){
                           if (x == "prop.err"){
                               x
                           } else {
                               paste0("SE.", substr(x, 2, nchar(x)))
                           }
                       }))
            .ret3 <- setNames(.ret$parFixedDf$`BSV(CV%)`,
                       sapply(rownames(.ret$parFixedDf),
                              function(x){
                           if (x == "prop.err"){
                               x
                           } else {
                               paste0("BSV.", substr(x, 2, nchar(x)))
                           }
                       }))
            .ret2 <- .ret2[names(.ret2) != "prop.err"]
            .ret3 <- .ret3[names(.ret3) != "prop.err"]
            .tmp <- c("VM", "KM", "Q", "Vp", "KA", "Cl")
            .tmp2 <- paste0("BSV.", .tmp)
            .tmp3 <- paste0("SE.", .tmp)
            .tmp <- c(.tmp, .tmp2, .tmp3)
            .ret <- data.frame(t(c(.ret1, .ret2, .ret3)), time=sum(.ret$time))
            for(.f in .tmp){
                if (is.null(.ret[[.f]])){
                    .ret[[.f]] <- NA
                }
            }
            .ret <- cbind(.ret, data.frame(t(.c)))
            return(.ret)
        } else {
            .out <- c("Cl", "Vc", "prop.err", "SE.Cl", "SE.Vc", "BSV.Cl", "BSV.Vc",  "VM", "KM", "Q", "Vp", "KA", "BSV.VM", "BSV.KM", "BSV.Q", "BSV.Vp",  "BSV.KA", "SE.VM", "SE.KM", "SE.Q", "SE.Vp", "SE.KA", "time");
            .ret <- cbind(data.frame(t(setNames(rep(NA, length(.out)), .out))),
                          data.frame(t(.c)))
            return(.ret)
        }
    })
    .w <- which(sapply(seq_along(.ret), function(x){!is.null(.ret[[x]])}));
    .ret <- .ret[.w]
    return(do.call("rbind", .ret))
}

```

# Setup general plot function

```{r}
library(ggplot2)
library(gridExtra)

xgx_theme_set()

f <- function(nlmixrVersion=c("1.1.1.3"), est=c("saem", "focei", "nlme"),
                     platform="unix"){
    ret <- readData(nlmixrVersion=nlmixrVersion, est=est,
                    platform=platform) %>%
        mutate(by=paste0(est, "(", ver, " ", os, ")"),
               run2=as.numeric(run))
    .lvl <- levels(ret$run)
    .brk <- seq_along(.lvl);

    .f <- function(var="Vc", tval=70, ylab="Vc (L)", bsvLab="BSV Vc(%)", bsvVal=30){
        p1 <- ggplot(ret, aes_string("run2", var, color="by")) +
            geom_point(size=3, alpha=0.5) +
            geom_line(alpha=0.5, size=1.3) +
            geom_hline(yintercept=tval, size=1.3) +
            theme(axis.text.x = element_text(face="bold", angle=45, hjust=1),
                  axis.title.y=element_text(face="bold", size=14),
                  axis.title.x=element_blank(),
                  legend.position="top") +
            labs(color="") +
            scale_x_continuous(breaks=.brk,labels=.lvl, minor_breaks=NULL) + 
            ylab(ylab)

        p2 <- ggplot(ret, aes_string("run2", paste0("BSV.", var), color="by")) +
            geom_point(size=3, alpha=0.5) +
            geom_line(alpha=0.5, size=1.3) +
            geom_hline(yintercept=bsvVal, size=1.3) +
            theme(axis.text.x = element_text(face="bold", angle=45, hjust=1),
                  axis.title.y=element_text(face="bold", size=14),
                  axis.title.x=element_blank(),
                  legend.position="none") +
            scale_x_continuous(breaks=.brk,labels=.lvl, minor_breaks=NULL) + 
            ylab(bsvLab)

        p3 <- ggplot(ret, aes_string("run2", paste0("SE.", var), color="by", group="by")) +
            geom_point(size=3, alpha=0.5) +
            geom_line(alpha=0.5, size=1.3) +
            theme(axis.text.x = element_text(face="bold", angle=45, hjust=1),
                  axis.title.y=element_text(face="bold", size=14),
                  axis.title.x=element_blank(),
                  legend.position="none") +
            scale_x_continuous(breaks=.brk, labels=.lvl, minor_breaks=NULL) + 
            ylab(paste0("SE log ", ylab))


        grid.arrange(p1, p2, p3, ncol=1)
    }
    suppressWarnings(.f())
    suppressWarnings(.f(var="Cl", tval=4, ylab="Cl (L/hr)", bsvLab="BSV Cl(%)"))
    suppressWarnings(.f(var="KA", tval=1, ylab="Ka (1/hr)", bsvLab="BSV Ka(%)"))
    suppressWarnings(.f(var="Vp", tval=40, ylab="Vp (L)", bsvLab="BSV Vp(%)"))
    suppressWarnings(.f(var="Q", tval=4.0, ylab="Vp (L/hr)", bsvLab="BSV Vp(%)"))
    suppressWarnings(.f(var="KM", tval=250, ylab="Km (mg/L)", bsvLab="BSV Km(%)"))
    suppressWarnings(.f(var="VM", tval=1000, ylab="Vmax (mg/hr)", bsvLab="BSV Vmax(%)"))
}

```

# Current Version plots
## SAEM, FOCEi and nlme

```{r}
nlmixrVersion <- as.character(packageVersion("nlmixr"));

f(nlmixrVersion, c("saem", "focei", "nlme"), "unix")
```
